# ORDINAL LOGISTIC REGRESSION USING THE `ordinal` PACKAGE

## **LOAD LIBRARIES**

```{r}
library(ggstatsplot)
library(ordinal)
library(pomcheckr)
library(gtsummary)
```

## **ATTACH DATA**

```{r}
df <- ordinal::wine
df$temp <- relevel(df$temp, ref="warm")
df$contact <- relevel(df$contact, ref="no")
attach(df)
View(df)
?ordinal::wine
```

## **EXPLORATORY & DESCRIPTIVE ANALYSIS**

```{r}
tbl_summary(df)

tcb <- with(df, temp:contact:bottle)[drop=T]
xtabs(response ~ tcb + judge, data = df)

with(df, {tcb <- temp:contact:bottle
  tcb <- tcb[drop=T]
  table(tcb, rating)})

grouped_ggbetweenstats(data=df, x=contact, y=response, grouping.var=temp, plot.type="box", type="nonparametric", pairwise.display="all")

grouped_ggbetweenstats(data=df, x=temp, y=response, grouping.var=contact, plot.type="box", type="nonparametric", pairwise.display="all")
```

## **ORDINAL REGRESSION ASSUMPTIONS**

1.  The dependent variable is measured on an ordinal level.

2.  Independent variables are either continious, categorical or ordinal.

3.  No multicollinearity - two or more independent variables should not be highly correlated with each other.

4.  Proportional Odds - each independent variable has an identical effect at each cumulative split of the ordinal dependent variable.

    ```{r}
    pomcheck(rating ~ temp*contact, data=df)
    plot(pomcheck(rating ~ temp*contact, data=df))
    ```

## **UNADJUSTED MODEL**

```{r}
fit1 <- clm(rating ~ temp + contact + temp*contact, data=df)
anova(fit1)
emmeans(fit1, pairwise ~ temp*contact)
grouped_ggbarstats(data=df, x=temp, y=rating, grouping.var=contact)
```

## ADJUSTED MODEL

```{r}
fit2 <- clm(rating ~ temp*contact + bottle + judge, data=df)
anova(fit2)
coef(fit2)
emmeans(fit2, pairwise ~ temp*contact)
```

## COMPARE MODELS

```{r}
AIC(fit1, fit2)
```
